Uncertainty propagation in neural networks for sparse coding

### 2018-11-29 nips_workshop_2018/camera_ready/BayesianLISTA.tex
updated camera-ready

 The rest of the paper is organised as follows. A NN approach for sparse coding is described in Section~\ref{subsec:nn_sc}. The Bayesian formulation is introduced in Section~\ref{subsec:bayesian_lista}. Section~\ref{sec:experiments} provides the experimental results. The proposed forward uncertainty propagation and probabilistic backpropagation methods are given in Appendices~\ref{sec:fprop} and~\ref{sec:backpropagation}. 
\section{Neural networks for sparse coding}
  \label{sec:bayesian_lista}
  The section first reviews existing NNs for sparse coding and then introduces the novel Bayesian NN.

\subsection{Frequentist neural networks}  
\label{subsec:nn_sc}
  The NN approach to sparse coding is based on earlier Iterative Shrinkage and Thresholding Algorithm (ISTA) \cite{daubechies2004iterative}. It addresses the sparse coding problem as the linear regression problem with the $l1$ penalty that promotes sparsity. For the linear regression model with observations $\mathbf{y} \in \mathbb{R}^K$, the design matrix $\mathbf{X} \in \mathbb{R}^{K \times D}$, and the sparse unknown vector of weights $\boldsymbol\beta \in \mathbb{R}^D$, ISTA minimises
  \begin{equation}
  \label{eq:regression_problem}
  ||\mathbf{X}\boldsymbol\beta - \mathbf{y}||_2^2 + \alpha ||\boldsymbol\beta||_1 \, \text{w.r.t.} \, \boldsymbol\beta,
  \end{equation}
where $\alpha$ is a regularisation parameter.

 \begin{wrapfigure}[13]{r}{0.5\textwidth}  
 \vspace{-10pt}
\begin{minipage}{0.5\textwidth}
  \begin{algorithm}[H]
    \caption{LISTA forward propagation}
\label{alg:lista}   
    \begin{algorithmic}[1]
      \REQUIRE observations $\mathbf{y}$, weights $\mathbf{W}, \mathbf{S}$, number of layers $L$
      \STATE Dense layer $\mathbf{b} \gets \mathbf{W}\mathbf{y}$ \label{eq:first_layer}
      \STATE Soft-thresholding function~$\widehat{\boldsymbol\beta}_0 \gets h_\lambda(\mathbf{b})$ \label{eq:thr_first}
      \FOR{$l=1$ \TO $L$}
        \STATE Dense layer $\mathbf{c}_l \gets \mathbf{b} + \mathbf{S}\widehat{\boldsymbol\beta}_{l-1}$ \label{eq:l_dense_layer}
        \STATE Soft-thresholding function $\widehat{\boldsymbol\beta}_{l} \gets h_\lambda(\mathbf{c}_l)$ \label{eq:l_thr}
      \ENDFOR
      \RETURN $\widehat{\boldsymbol\beta} \gets \widehat{\boldsymbol\beta}_{L}$
    \end{algorithmic}
  \end{algorithm}
\end{minipage}
\end{wrapfigure}  
At every iteration $l$, ISTA obtains the new estimate $\widehat{\boldsymbol\beta}_l$ of the target vector $\boldsymbol\beta$ as the linear transformation $\mathbf{b} = \mathbf{W}\mathbf{y} + \mathbf{S}\widehat{\boldsymbol\beta}_{l-1}$ propagated through the soft-thresholding function %$h_\lambda(\cdot)$
  \begin{equation}
  h_\lambda(\mathbf{b}) = \text{sgn}(\mathbf{b}) \max(|\mathbf{b}| - \lambda, 0),
  \end{equation}
  where $\lambda$ is a shrinkage parameter.
  In ISTA, weights $\mathbf{W}$ and $\mathbf{S}$ of the linear transformation are assumed fixed.

@@

 Appendix provides additional results on predictive performance on the MNIST data.

@@

\section{Appendix: Predictive performance on MNIST data}
In this experiment, the methods are evaluated on the MNIST dataset in terms of predictive performance. We use $100$ images for training and $100$ for test.

>>>>

 The rest of the paper is organised as follows. A NN approach for sparse coding is described in Section~\ref{subsec:nn_sc}. The Bayesian formulation is introduced in Section~\ref{subsec:bayesian_lista}. Section~\ref{sec:experiments} provides the experimental results. The proposed forward uncertainty propagation and probabilistic backpropagation methods are given in Appendices~\ref{sec:fprop} and~\ref{sec:backpropagation}. 
\section{Neural networks for sparse coding}
  \label{sec:bayesian_lista}
  This section presents background knowledge about networks for sparse coding and then describes the novel Bayesian neural network. 
\subsection{Frequentist neural networks}  
\label{subsec:nn_sc}
  The NN approach to sparse coding is based on earlier Iterative Shrinkage and Thresholding Algorithm (ISTA) \cite{daubechies2004iterative}. It addresses the sparse coding problem as the linear regression problem with the $l1$ penalty that promotes sparsity. For the linear regression model with observations $\mathbf{y} \in \mathbb{R}^K$, the design matrix $\mathbf{X} \in \mathbb{R}^{K \times D}$, and the sparse unknown vector of weights $\boldsymbol\beta \in \mathbb{R}^D$, ISTA minimises
  \begin{equation}
  \label{eq:regression_problem}
  ||\mathbf{X}\boldsymbol\beta - \mathbf{y}||_2^2 + \alpha ||\boldsymbol\beta||_1 \, \text{w.r.t.} \, \boldsymbol\beta,
  \end{equation}
where $\alpha$ is a regularisation parameter.

 \begin{wrapfigure}[13]{r}{0.5\textwidth}  
 \vspace{-25pt}
\begin{minipage}{0.5\textwidth}
  \begin{algorithm}[H]
    \caption{LISTA forward propagation}
\label{alg:lista}   
    \begin{algorithmic}[1]
      \REQUIRE observations $\mathbf{y}$, weights $\mathbf{W}, \mathbf{S}$, number of layers $L$
      \STATE Dense layer $\mathbf{b} \gets \mathbf{W}\mathbf{y}$ \label{eq:first_layer}
      \STATE Soft-thresholding function~$\widehat{\boldsymbol\beta}_0 \gets h_\lambda(\mathbf{b})$ \label{eq:thr_first}
      \FOR{$l=1$ \TO $L$}
        \STATE Dense layer $\mathbf{c}_l \gets \mathbf{b} + \mathbf{S}\widehat{\boldsymbol\beta}_{l-1}$ \label{eq:l_dense_layer}
        \STATE Soft-thresholding function $\widehat{\boldsymbol\beta}_{l} \gets h_\lambda(\mathbf{c}_l)$ \label{eq:l_thr}
      \ENDFOR
      \RETURN $\widehat{\boldsymbol\beta} \gets \widehat{\boldsymbol\beta}_{L}$
    \end{algorithmic}
  \end{algorithm}
\end{minipage}
\end{wrapfigure}  
At every iteration $l$, ISTA obtains the new estimate $\widehat{\boldsymbol\beta}_l$ of the target vector $\boldsymbol\beta$ as the linear transformation $\mathbf{b} = \mathbf{W}\mathbf{y} + \mathbf{S}\widehat{\boldsymbol\beta}_{l-1}$ propagated through the soft-thresholding function %$h_\lambda(\cdot)$
  \begin{equation}
  h_\lambda(\mathbf{b}) = \text{sgn}(\mathbf{b}) \max(|\mathbf{b}| - \lambda, 0),
  \end{equation}
  where $\lambda$ is a shrinkage parameter.
  In ISTA, weights $\mathbf{W}$ and $\mathbf{S}$ of the linear transformation are assumed fixed.

@@

 Appendix~\ref{app:mnist} provides additional results on predictive performance on the MNIST data.

@@

\section{Appendix: Predictive performance on MNIST data}
\label{app:mnist}
In this experiment, the methods are evaluated on the MNIST dataset in terms of predictive performance. We use $100$ images for training and $100$ for test.
